\textit{Supplemental Information:}\\ Design Principles from Glass Sponges \\for Structurally Robust Lattices

### 2018-10-24 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
new version

\documentclass[10pt,twocolumn,twoside]{fernandes_paper} 
%Default opts:9pt,twocolumn,twoside
\graphicspath{{images_paper/}}

\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}

\title{Utilizing Design Principles from Glass Sponges for Structurally Robust Lattices}

@@

\keywords{Architected materials, truss structure, buckling resistance, glass sponges, bio-inspired engineering.}

\usepackage{comment}

\begin{document}
\maketitle
% \doublespace
%\linenumbers

\begin{figure*}[ht]
  \captionsetup{width=0.8\textwidth}
\begin{center}
\includegraphics[width=0.9\textwidth]{Fig1}
\end{center}
\caption{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge. (c) Close up microscope image of the sponge. (d) Comparison between the idealized model (green and blue lines) and the sponge structure. (e) Representative volume element unit cell of the idealized model.}\label{Fig1}
\end{figure*}

\begin{figure*}[ht]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.9\textwidth]{Fig2}
	\caption{\textbf{Experimental Results.} (a) shows the different designs (Design A-D) considered in this analysis. (b) shows the undeformed experimental setup for the different designs. (c) shoes the post-buckling deformed state for each of the designs. (d) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.}\label{Fig2}
\end{figure*}

\begin{figure*}[ht]
	\captionsetup{width=0.8\textwidth}
	\begin{center}
		\includegraphics[width=0.7\textwidth]{Fig3}
	\end{center}
	\caption{\textbf{Structure Mechanical Response} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs. The designs considered are depicted below as Design A-D.} \label{Fig3}
\end{figure*}

\begin{figure*}[ht]
	\captionsetup{width=0.8\textwidth}
	\begin{center}
		\includegraphics[width=0.7\textwidth]{Fig4}
	\end{center}
	\caption{\textbf{Critical Buckling Load Optimization Results.} (a) shows the optimal value of critical buckling load for varying number of diagonals. The color of each point represents the optimal mass ratio $\lambda$  parameter for that configuration.  Note, we do not explicitly show the other optimal values for the diagonal spacing parameters. (b) shows the resulting deformed geometries for designs including one to four diagonals.} \label{Fig4}
\end{figure*}

The design of diagonal reinforcement for truss systems found in bridges, buildings, and beam structures has not changed much since the introduction of the Long truss design patented by Colonel Long in 1830 \citep{waddell1916}. In modern day structures, we see many different iterations of new truss designs with various applications in mind but with what is the same fundamentally simple diagonal reinforcement technique. This technique, involving two diagonal trusses (one in each direction) crossing the opening of a square beam cell, limits the transverse motion of the main support beams ultimately giving the structure shear capacity and strength. Alternative, methods to strengthen beam structures have been introduced via material \hl{[...],} \mf{needs more information and citations}. However, in this study we seek to further strengthen square lattices through inspiration from a particular sponge species called \textit{Euplectella Aspergillum}, commonly known as the Venus' flower basket. This particular species, which is part of the \textit{Hexactinellid} class, is primarily made of silica and exhibit a periodic and regular grid structure with an intrinsic and different diagonal reinforcement scheme.

\textit{Hexactinellids}, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m. Beyond their fracture propagation inhibiting material composition \citep{weaver2007}, they are perceived to exhibit large structural rigidity and strength against buckling.

Structurally, they are composed of a base square-grid architecture and regular ordering of vertical and horizontal struts that form the skeletal system. Furthermore, their base structure is overlaid with double diagonal (two diagonals in each direction) reinforcement struts, which create a checkerboard-like pattern of open-closed cell structure. This diagonal reinforcement design is conjectured to give the sponge greater buckling resistance and strength to localized damage then it would experience having a single diagonal reinforcement strut (with equivalent allocation of mass between the diagonal reinforcements.)

The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?}

In this paper, we present a numerical and experimental mechanical analysis of multiple  structures under various loading conditions as well as survey different arrangements within similar design space via parametric studies. Furthermore, we incorporate an optimization algorithm on our parameter space to obtain an optimal design to withstand (via critical buckling load) uniaxial loading parallel to the non-diagonal truss members. For this analysis, we restrict ourselves to four designs considered (Design A-D) that are shown along the top of \cref{Fig2}. As part of our analysis, we study the critical buckling strain and the elastic load carrying capacity of these structures.

Numerically, we compare the linear stiffness and critical buckling strain for each of the different designs using Finite Element Analysis (FEA). To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI). For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections. In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.

The results presented in \cref{Fig3}(a) show that all of the structures containing diagonal reinforcement behave the same way for varying loading angles. This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to the perpendicular aligned beam elements. This is further supported by the graph for Design D, where all of the material from the diagonal beams is removed and allocated to the non-diagonal elements, giving the structure a higher stiffness at $0^\circ$ and lower stiffness at $45^\circ$. Any stiffness at $45^\circ$ in Design D is strictly a result of the bending resistance due to the pinned joints between the non-diagonal beams.

The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements. From this figure, we can see that Design A, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement. At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of Design C, which is the commonly used truss systems.

For the experimental validation, we conducted various uniaxial compression tests of 3D printed planar beam structures (with circular cross-section) on an Instron 5566. The specimen was 3D printed using \hl{xxxx} material with manufacturer reported \hl{xxxx GPa} modulus. The printed specimen consisted of a 10x10 cell structure spanning a total of 10x10 inch planar structure. The experimental set-up consisted of two \hl{1 inch} thickness acrylic sheets bolted together along the perimeter and spaced from each other by the thickness of the 3D printed specimen.\\\\\hl{[...]} \mf{more to go here when we have experimental setup and data}

In order to study the impact of diagonal separation for Design A (sponge design), we perform a parametric study on the buckling behavior of the structure under varying separation distances ($C$). Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles. {Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$). Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.

To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements. To quantitatively describe this mass ratio, we introduce the variable $\lambda$ as the fraction $V_{nd}/V_d$, where $V_{nd}$ describes the collective volume of all non-diagonal elements and $V_d$ is the collective volume of all diagonal elements in the unit cell. It is important to note, that for all simulations varying $\lambda$ we maintain the same total mass of the structure, namely $V_{nd}+V_d=\mbox{const. } \forall (V_{nd},V_d)$.

%\section*{Acknowledgments}

% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship
% Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft

\nocite{*}
\bibliography{refs}
\bibliographystyle{apalike}

>>>>

\documentclass[9pt,twocolumn,twoside]{fernandes_paper} 
%Default opts:9pt,twocolumn,twoside
\graphicspath{{images_paper/}}

\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
% \newcommand{\mf}[1]{\noindent\color{color2}\sethlcolor{cyan}\hl{$\Longrightarrow$#1}\normalcolor}
\newcommand{\mf}[1]{\noindent\color{orange}{$\Longrightarrow$#1}\normalcolor}

\title{Design Principles from Glass Sponges \\for Structurally Robust Lattices}

@@

\keywords{Architected materials, truss structure, square lattice, buckling resistance, glass sponges, bio-inspired engineering.}

\begin{document}
\maketitle
% \doublespace
\linenumbers

\begin{figure*}[!htb]
    \centering
    \captionsetup{width=0.8\textwidth}
    \includegraphics[width=0.6\textwidth]{Fig1}
    \caption{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge. (c) Close up microscope image of the sponge. (d) Comparison between the idealized model (green and blue lines) and the sponge structure.}\label{Fig1}
\end{figure*}

\begin{figure*}[!htb]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.9\textwidth]{Fig2}
	\caption{\textbf{Experimental Results.} (a)-(d) shows the different designs (\emph{Design A-D}) considered in this analysis. (e) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.(f) shows the undeformed experimental setup for the different designs.}\label{Fig2}
\end{figure*}

The hexactinellida sponge \textit{Euplectella Asperguillum} species, commonly known as the Venus' Flower Basket, has received significant attention in the engineering and material science communities for its remarkable hierarchical design and robustness across many scales. Its fundamental building blocks, called spicules, are constructed using laminated skeletal elements which consist of a central protein-based organic core surrounded by alternating concentric domains of consolidated silica nanoparticles (clear glass) and organic interlayers. \cite{weaver2010,aizenberg2005,shulze1887} Remarkably, these spicules are arranged to form a regular and complex square-grid architecture reinforced by a set of double diagonals (a set of two diagonals going in each direction), creating a checkerboard-like pattern of alternating open-closed cells (see \cref{Fig1}). Notably, while it has been demonstrated that the laminar architecture of the spicule leads to superior resistance to fracture propagation \cite{miserez2008} and the shape of the spicules enhances the overall strength against buckling \cite{monn2015}, the functionality of the complex double diagonal square-grid macroscale structure is still unknown.

The macro-scale architecture of the \textit{Euplectella Asperguillum} provides a remarkable example of slender and regular lattice-like material constructed by nature. Such lattice materials have recently demonstrated to offer novel and unique properties, including being light weight, \cite{schaedler2011,ashby2006} having high energy absorption, \cite{evans2001} and having the ability to control the propagation of waves \cite{phani2006} as well as heat flow. \cite{lu1998,ashby1997,evans1998} Generally speaking, the properties and functionality of such materials are  dictated by the node connectivity. In particular, a minimum node connectivity of 6 is required for 2D lattice materials to be  stretching-dominated, and therefore,  more weight-efficient for structural applications. \cite{deshpande2001} As such, lattice materials with square architecture (which has a node connectivity of 4) are known to be very unstable when loaded along specific directions (as the only shear resistance arises from the joints) \cite{gibson1999} and typically require diagonal bracings to stabilize them.

In this paper, we use the anatomy of the hexactinellid sponge species \textit{Euplectella Aspergilum} skeletal system as inspiration for the design of mechanically robust lattice materials with square architecture. First, we use a combination of experiments and numerical analyses to understand the overarching mechanical functions driving the particular and regular alignment of trusses found in the sponge. Then, we utilize an optimization algorithm based on the principles introduced by the sponge to determine the  configuration of reinforced square lattice that achieve the highest mechanical performance. Therefore, our results indicate that the  principles used by Nature to design the \textit{Euplectella Aspergilum}  may guide the design of lattice architectures structurally more robust than those currently used in modern devices and infrastructure.

To understand  the role of the sponge architecture, we start by comparing its performance to that of other three 2D lattices, all based on a   square architecture with edges of length $L$ (see Supplementary Materials for details). Specifically,  we  consider:
\emph{Design A}  which is inspired by the sponge and comprises a set of double diagonals going in opposite directions (see \cref{Fig2}(a)); \emph{Design B}  which is similar to the sponge design but only contains a single diagonal crossing each of the closed cells (see \cref{Fig2}(b));   \emph{Design C}  which is similar to bracings found in modern structural engineering and contains a crossing set of diagonal beams in every cell (see \cref{Fig2}(c)); and \emph{Design D}  which  has no diagonal reinforcement (see \cref{Fig2}(d)). In all our designs we consider struts with rectangular cross section and depth large enough to avoid any out-of-plane deformation. Moreover, to ensure a fair mechanical comparison between the four designs, we keep the total mass constant and
for \emph{Designs A, B}, and \emph{C} we also constrain the  ratio between the mass of the diagonals and non-diagonals struts to be identical (see Supplementary Materials for details).  %Finally, in all our analyses we constraint the lattices to deform in-plane. To this end, in our experiments we consider specimens sufficiently thick for which out-of-plane deformation is negligible and in the the numerical simulations we assume plane-strain conditions.

We started by testing samples of the four considered designs comprising
a tessellation of $6\times6$ square cells with $L=$ \KB{??} mm. The samples were fabricated with a multi-material \mf{@Katia: what do you mean multi-material} 3D printer \mf{(add model @James)} using \KB{name of material} \mf{add info here @James} and compressed uniaxially  using  an Instron Model 5969 with 50kN capacity  (\Cref{Fig2}(e)). \KB{switch labels (f) and (e) in Fig 2} %Using this set-up, we tested three replicas of each design that were 3D printed and tested on different days.
The results reported in \Cref{Fig2}(f)  show that all structures are characterized by an initial linear behavior, a sudden departure from linearity caused by in-plane buckling and catastrophic failure following immediately thereafter.   Furthermore, it is also important to note that all  designs with diagonal reinforcement (i.e. \emph{Designs A-C}) overlay each other in the linear regime. Focusing on the initial linear regime, we find that  the different diagonal reinforcement designs considered do not impact the structure's overall stiffness (i.e. the initial stiffness of Designs A, B and C is identical). Design D, as expected, has a higher stiffness because of its thicker vertical and horizontal element's. However, it fails much earlier due to a shorter critical buckling strain. On the other hand, \emph{Design A} (the sponge design) outperforms the  considered structures in the amount of load it can bare before the onset of buckling occurs and provides the best mechanical performance without compromising the structural stiffness.

Next, we use Finite Element (FE) simulations conducted using the commercial software ABAQUS/Standard to further understand how the double diagonal system found in the \textit{Euplectella Aspergilum} results in significantly improved mechanical performance.  All our models are constructed using  Timoshenko beam elements (Abaqus element type \KB{B??}) and the material's response is captured using an isotropic linear elastic materials. We start by simulating the response of the  6 by 6 specimens tested in our experiments and find  very good agreement  between our experiments and numerical model (see Fig. \KB{??}).  Having validated our numerical analysis, we next use them to investigate \KB{what??? big picture of Fig 3} the To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI). For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections. In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.

The results presented in \cref{Fig3}(a) show that all of the structures containing diagonal reinforcement behave the same way for varying loading angles. This illustrates the fact that the structural stiffness is independent of design, but instead depend on the amount of material allocated to the perpendicular aligned beam elements. This is further supported by the graph for \emph{Design D}, where all of the material from the diagonal beams is removed and allocated to the non-diagonal elements, giving the structure a higher stiffness at $0^\circ$ and lower stiffness at $45^\circ$. Any stiffness at $45^\circ$ in Design D is strictly a result of the bending resistance due to the pinned joints between the non-diagonal beams.

The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements. From this figure, we can see that \emph{Design A}, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement. At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of \emph{Design C}, which is the commonly used truss systems.

%In order to study the impact of diagonal separation for Design A (sponge design), we perform a parametric study on the buckling behavior of the structure under varying separation distances ($C$). Note, we do not perform a stiffness test as the linear stiffness is not dependent on design but instead depends on the mass ratio between diagonal and non-diagonal elements as well as the diagonal element's angles. {Fig5}(a) shows the buckling behavior as a function of the non-dimensional separation ($C/L$). Where at $C/L=0$ we obtain Design B, and at $C/L=0.5$ we obtain diamond shaped reinforcement alignment.

%To investigate the effect of mass ratio on the properties of the structure, we parametrize the allocation of mass between the diagonal elements and the non-diagonal elements. To quantitatively describe this mass ratio, we introduce the variable $\lambda$ as the fraction $V_{nd}/V_d$, where $V_{nd}$ describes the collective volume of all non-diagonal elements and $V_d$ is the collective volume of all diagonal elements in the unit cell. It is important to note, that for all simulations varying $\lambda$ we maintain the same total mass of the structure, namely $V_{nd}+V_d=\mbox{const. } \forall (V_{nd},V_d)$.

\mf{write section on Optimization here}

In summary, the investigation of the \textit{Euplectella Asperguillum's} intricate and regular structure has lead to macro-scale design principles that can be applied to general square lattice materials and structures. When compared with other diagonally reinforced square lattice designs, the sponge design has experimentally and numerically proven to provide a superior mechanism for withstanding loads prior to onset of buckling. By controlling the number of diagonals, diagonal separation and mass ratio between diagonal and non-diagonal elements, a structure's design can be optimized to withstand buckling load caring capacity.

\section*{Acknowledgments}
 \noindent This work was partially supported by: NSF-GRFP Fellowship  Grant\# DGE-1144152 (for M. C. F.), GEM Consortium Fellowship (for M. C. F.), the Harvard Graduate Prize Fellowship (for M. C. F.), \mf{add whatever else here}.
 We would also like to thank James R. Rice, Mark Scott, Joost Vlassak, and Chris Rycroft for insightful discussions.

\begin{figure*}[!htb]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.7\textwidth]{Fig3}
	\caption{\textbf{Structure Mechanical Response} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs. The designs considered are depicted below as \emph{Design A-D.}} \label{Fig3}
\end{figure*}

\begin{figure*}[!htb]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.7\textwidth]{Fig4}
	\caption{\textbf{Critical Buckling Load Optimization Results.} (a) shows the optimal value of critical buckling load for varying number of diagonals. The color of each point represents the optimal mass ratio $\lambda$  parameter for that configuration.  Note, we do not explicitly show the other optimal values for the diagonal spacing parameters. (b) shows the resulting deformed geometries for designs including one to four diagonals.} \label{Fig4}
\end{figure*}

\noindent\rule[-1.5ex]{\linewidth}{1pt}
%\nocite{*}
\bibliography{refs}
%\bibliographystyle{apalike}
%\bibliographystyle{plainnat}
\bibliographystyle{unsrtnat}

### 2018-10-24 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
new version

\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}

\title{Supplemental Information:\\ Harnessing  Design Principles from Glass Sponges for Structurally Robust Lattice Structures}

@@

\begin{document}
\maketitle
% \doublespace
% \linenumbers
\section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
The periodic structures investigated in this study are inspired by the skeleton of the Hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket". In this section we provide a detailed description of the sponge geometry and measured dimensions.

@@

@@

Finally, it is important to note that in this study we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with circular cross section to be equal to that of Design A, namely,
\begin{equation}\label{con1}
V_{\alpha,d}+V_{\alpha,nd}={V}_{A,T}=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right),
\end{equation}
with $\alpha=$ B, C and D.
For Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A
\begin{equation}\label{con2}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\frac{V_{A,nd}}{V_{A,d}}=2\sqrt{2},
\end{equation}
with $\alpha=$ B and C.

@@

Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely,
\begin{equation}\label{con1R}
{V}_{A,T}=V_{\alpha,d}+V_{\alpha,nd}= 8LH(T_{A,nd}+\sqrt{2}T_{A,d})=8LHT_{A,nd}\left(1+\frac{1}{\sqrt{2}}\right),
\end{equation}
with $\alpha=$ B, C and D.
Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A,
\begin{equation}\label{con2R}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\sqrt{2},
\end{equation}
with $\alpha\in$ B and C.

@@

\begin{figure}
    \centering
	\includegraphics[width=0.6\linewidth]{SFig10.png}
	\caption{{\bf RVE used for the different designs.} This figure portrays the different RVEs used for the different designs considered. Each unit cell is shifted by half of a square cell to the horizontally and half of a square cell  vertically. Periodic boundary conditions are applied on the nodes that intersect with the light gray dashed line.}
	\label{RVE}
\end{figure}

@@

@@

\section{Local and Global Instabilities}
\mf{How should I go about describing this here?}
\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{SFig12.pdf}
	\caption{{\bf Global vs Local Instabilities.} This figure shows four identical analysis performed on the Designs A-D. For each analysis we obtain the critical buckling strain for varying $M\times M$ tessellations (y-axis) of the unit cell also for varying loading angle $\theta$ (x-axis). The color of a pixel corresponds to a value of critical buckling strain for that simulation. For each of the simulations, periodic boundary conditions are applied along the outer perimeter of the $M\times M$ structure. This plot conveys that for Designs A-C the prominent buckling mode is the local mode, whereas for Design D, the prominent mode is a global mode. Choosing a sufficiently large $M$ allows Design D to converge to a finite value for each $\theta$.}
	\label{DomainSize}
\end{figure}
% Bibliography
%\nocite{aizenberg2005}
%\nocite{deshpande2001}
%\nocite{miserez2008}
%\nocite{weaver2010}

\bibliography{refs}
\bibliographystyle{apalike}
% \bibliographystyle{plainnat}

>>>>

\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
\newcommand{\mf}[1]{\noindent\color{color2}\sethlcolor{cyan}\hl{$\Longrightarrow$#1}\normalcolor}

\title{\textit{Supplemental Information:}\\ Design Principles from Glass Sponges \\for Structurally Robust Lattices}

@@

\begin{document}
\maketitle
% \doublespace
\linenumbers

\section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
The periodic structures investigated in this study are inspired by the skeleton of the Hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket". In this section we provide a detailed description of the sponge geometry and measured dimensions.

@@

\subsection{Image Segmentation Analysis}
\mf{@James, add details of image segmentation analysis performed for extracting diameter rations between diagonals and non-diagonals}

@@

Finally, it is important to note that in this study we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with circular cross section to be equal to that of Design A, namely,
\begin{equation}\label{con1}
V_{\alpha,d}+V_{\alpha,nd}={V}_{A,T}=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right),
\end{equation}
with $\alpha=$ B, C and D.
For Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A
\begin{equation}\label{con2}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\frac{V_{A,nd}}{V_{A,d}}=2\sqrt{2},
\end{equation}
with $\alpha\in$ B and C.

@@

Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely,
\begin{equation}\label{con1R}
{V}_{A,T}=V_{\alpha,d}+V_{\alpha,nd}= 8LH(T_{A,nd}+\sqrt{2}T_{A,d})=8LHT_{A,nd}\left(1+\frac{1}{\sqrt{2}}\right),
\end{equation}
with $\alpha\in$ B, C and D.
Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A,
\begin{equation}\label{con2R}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\sqrt{2},
\end{equation}
with $\alpha\in$ B and C.

@@

\begin{figure}
    \centering
	\includegraphics[width=0.9\linewidth]{SFig10.pdf}
	\caption{{\bf RVE used for the different designs.} This figure portrays the different RVEs used for the different designs considered. Each unit cell is shifted by half of a square cell to the horizontally and half of a square cell  vertically. Periodic boundary conditions are applied on the nodes that intersect with the light gray dashed line.}
	\label{RVE}
\end{figure}

\section{Experimental Setup}
\mf{add full description of experimental setup and full table of 3D printed material properties; @James: I need to material properties}

@@

\mf{Add table with optimal parameters results for each N, for both circular and rectangular cross-sections}

@@

\section{Local and Global Instabilities}
\mf{How should I go about describing this here?}
\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{SFig12.pdf}
	\caption{{\bf Global vs Local Instabilities.} This figure shows four identical analysis performed on the Designs A-D. For each analysis we obtain the critical buckling strain for varying $M\times M$ tessellations (y-axis) of the unit cell also for varying loading angle $\theta$ (x-axis). The color of a pixel corresponds to a value of critical buckling strain for that simulation. For each of the simulations, periodic boundary conditions are applied along the outer perimeter of the $M\times M$ structure. This plot conveys that for Designs A-C the prominent buckling mode is the local mode, whereas for Design D, the prominent mode is a global mode. Choosing a sufficiently large $M$ allows Design D to converge to a finite value for each $\theta$.}
	\label{DomainSize}
\end{figure}

\section{Stress Concentration}
\mf{add section on stress concentration for plane strain setting}

% Bibliography
\bibliography{refs}
%\bibliographystyle{apalike}
% \bibliographystyle{plainnat}
\bibliographystyle{unsrtnat}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{SFig13.pdf}
	\caption{{\bf asdf.} asdf}
	\label{asdfa}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{SFig14.pdf}
	\caption{{\bf asdf.} asdf}
	\label{asdfgas}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{SFig15.pdf}
	\caption{{\bf asdf.} asdf}
	\label{asdfgdfsgas}
\end{figure}
